---
title: "Assignment 5 - K-means"
author: "Nikita Bhusal"  
format:
  html:
    code-fold: true 
    embed-resources: true
    toc: true
    number-sections: true
    theme: cerulean
editor_options: 
  chunk_output_type: inline
--- 

# Goal of this assignment  
The goal of this assignment is for you to **explore the k-means algorithm** learned in class.  

In specific, you will be asked to:  
  - Use a different number of k than what we did in class,  
  - Validate the results of the new model, and  
  - Compare your results with those found in class  

# Instructions  
You will first need to watch (and code along with) the following k-means video:

  - k-means lab: https://youtu.be/GFgMp5tYiMU?si=CI0E-2r-cYZLGVL1 (start from beginning of video, watch till the 01:10)  
  - The partial script for the video above is in our GitHub repository, 04-classcode, **02-24_multivar_kmeans_partial.qmd**. Move the script mentioned above into the `code` subfolder of your `06_multivar` project
  
**AFTER watching the video and developing the code in class**, then start working on this HW assignment (instructions below):
  - Move this script into the `code` subfolder of your `06_multivar` project  
  - On this script, fill your name under the `title` section of the YAML  
  - Go back to the class k-means script, choose a different number of k than what we did in class, and run it.    


```{r}

library(tidyverse)
library(ggcorrplot)
library(broom)
library(car)
library(factoextra)
library(ggpmisc)
```


```{r}
weather <- read.csv("../data/weather_monthsum.csv")

weather
```

Since k-means only take numerical variables, let's select them now.  
```{r selecting only numerical vars}
weather_n <- weather %>%
  dplyr::select(-c(year:strength_gtex))

weather_n
```


```{r}
# normalizing the data
weather_norm <- weather_n %>%
  mutate(across(everything(),~scale(.x)))

weather_norm
summary(weather_norm)
```

Since the choice of k can be subjective, we will need to find an objective way to select the value of k that most properly represents our dataset.  

```{r choosing k - total error}
# Total error x k
fviz_nbclust(weather_norm,
             method = "wss",
             k.max = 10,
             FUNcluster = kmeans)
```




# Questions  
## 1. What number of k did you select? Why?  


The reasons to select k = 3 are: 

1. The total error is 3 to 4 from the above. Although the error shows 3-4, the large drop or elbow started from k=3. 
2. With k = 4, cluster 2 has comparatively less number of observation (16) so  it might be good to check with reducing the clusters to 3. 
 

```{r mod_km3 }
mod_km3 <-kmeans(weather_norm,
                 centers = 3,
                 nstart = 10,
                   )

mod_km3

```

## 2. How many observations are there in each cluster?  

```{r}
weather %>%
  mutate(cluster = mod_km3$cluster) %>%
  group_by(cluster) %>%
  tally()
```


The cluster 1 has the highest no. of observations; 337, followed by cluster 2 with n= 253 and cluster 3 with 108. 




## 3. Using the `fviz_cluster()` function, import here the original plot with k=4 did in class, and the new plot with the number of k you selected for this exercise. How do they visually compare? Which one seems to be a better choice, and why?  

```{r mod_km4 }
 
## k = 4 (From Class) to compare k=4 and k=3

mod_km4 <-kmeans(weather_norm,
                 centers = 4,
                 nstart = 10,
                   )

mod_km4

## K means = 4 PCA plot

fviz_cluster(mod_km4,
             data = weather_norm
             )
```


```{r r kmeans_k3_pca_plot}
fviz_cluster(mod_km3,
             data = weather_norm
             )
```
After comparing both plots with k= 4 and k= 3, the plot with k=3 looks better. There are few reasons for this:
1. From the elbow plot, the large drop started from k=3 as there is a clear bend at this point.
2. Visually, the clusters in k=3 shows well-separated and compact clusters compared to the clusters from k=4. At 4-clusters, the is a small fraction of cluster seen on the lower-left hand side which does not substantially separate the grouping or possible extra-cluster. Therefore for better visualization and interpretation, k=3 is more appropriate choice. 


# Submitting your work  
Once you have developed all the code and answers, make sure to Render this quarto file.  

**Notes on rendering**:  

- Make sure to render your work and inspect how the final html look like.  
- If it does not look professional for whatever reason, then fix the issue, re-render it, recheck.  
- Only send me your work once your html file looks professional.  
  - **DO NOT** delete the file's heading levels (# and ##). They set up the proper heading 1 and 2 levels, and I use them to guide my grading.  
  - If a given chunk is also outputting warnings or messages, inhibit this behavior by changing the chunk options `message` and `warning` to `FALSE` (I'll demo this in a moment).  
  
  - If, after rendered, 2 lines of text are connected and you wish to "break line" between them, add 2 extra spaces after the first one (I'lld demo this in a moment).  

After rendering, an .html file will be created on your `code` folder.  

Rename this file to `Assignment-05-kmeans-LASTNAME.html`.    

For ex., mine would be `Assignment-05-kmeans-Bastos.html`.

Submit your work by uploading the **html** file to **eLC Assignment #5 - K-means** by Monday March 2nd 11:59 pm.  
  




